On exponential sums arising from the classical doubling method

### 2025-04-11 main.tex
(back to default color scheme)
I edit a lot of times in black background, write text
Typos

The main result of this paper is 
\begin{theorem}
\label{thm:Phi_stable}
	The assignments $H \to \cComplex$ given by $$h \mapsto \genHermitianJacobiKernel{\hermitianSpace}{\chi}\left(h\right)$$ are stable functions. Equivalently, the assignments $$\pi \mapsto \dblJacobiSumScalar{\pi}{\chi}$$ are constant on geometric Lusztig series.
\end{theorem}
This result will be proved in \Cref{subsec:stability}. Our next goal is to prove the following intermediate result.
\begin{lemma}
\label{lem:indep_geo_conj}
    If $(\algebraicGroup{T}_H,\theta_H)$ and $(\algebraicGroup{T}'_H, \theta_H)$ are geometrically conjugate, then $$\frac{\dblVirtualJacobiSumScalar{\RTTheta{T_H}{\theta_H}}{\chi}}{\RTTheta{T_H}{\theta}(1)} = \frac{\dblVirtualJacobiSumScalar{\RTTheta{T'_H}{\theta'}}{\chi}}{\RTTheta{T'_H}{\theta'_H}\left(1\right)}.$$
\end{lemma}

We recall that \elad{Maybe we can help the reader and tell them that $m=1,2$} $$
	\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N} = \prod_{a \in \mathbb{Z}/m\mathbb{Z}} \left(\algebraicGroup{\GL}_N\right)_a,
$$
with Frobenius root action $$\Frobenius(g_0,g_1, \hdots, g_{m-1}) = (\Frobenius(g_{m-1}), \Frobenius(g_0), \Frobenius(g_1), \hdots, \Frobenius(g_{m-2})).$$ We will freely identify $\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N}(\finiteField) = \GL_N(\finiteFieldExtension{m})$ via the map $\left(g_0,\dots,g_{m-1}\right) \mapsto g_0$.

@@

%Given a rational maximal torus $\algebraicGroup{T} \subset \restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n}$, a character of $\algebraicGroup{T}(\mathbb{F})$, and a non-trivial additive character $\psi$ of $\finiteFieldExtension{m}$, we define a Guass sum
%\[
%	\tau_{\algebraicGroup{T}}(\theta,\psi) := \varepsilon_{\algebraicGroup{
%    \algebraicGroup{T}}}\sum_{t\in \tilde{\algebraicGroup{T}}^{\Frobenius}} \theta^{-1}(t)\psi(\trace(t)).
%\]
%\calvin{maybe should do inverses in the definition. Not sure if $\tau$ or $g$ make more sense.}

Given a torus character pair $(\algebraicGroup{T},\theta)$ with associated $\Frobenius$-twisted conjugacy class $(w) \in W(\algebraicGroup{T}_0)$ along with an embedding $\rho:\algebraicGroup{G}^* \to \restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N}$ and a Frobenius stable maximal torus $\algebraicGroup{T}^* \subset \restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N}$ such that $\rho(\algebraicGroup{T}_0^*) \subset \algebraicGroup{T}^*$, let $\tilde{\algebraicGroup{T}}$ be the torus of $\algebraicGroup{\GL}_N$ \elad{of $\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N}$?} corresponding to the $\Frobenius$-twisted conjugacy class of $\rho(w) \in W(\algebraicGroup{T})$ \elad{of $W\left(\algebraicGroup{T}^{\ast}\right)$? Or maybe $W\left(\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{A}_N}\right)$?} and let $\tilde{\theta} \colon \tilde{\algebraicGroup{T}}^{\Frobenius} \to \multiplicativegroup{\cComplex}$ be a character such that $(\rho(s_{\algebraicGroup{T},\theta})) = (s_{\tilde{\algebraicGroup{T}},\tilde{\theta}})$. We define
\[
    g_{\algebraicGroup{T},\rho}(\chi,\theta,\psi) := \tau_{\Tilde{\algebraicGroup{T}}}(\tilde{\theta} \times \chi \circ \det,\psi) = \epsilon_{\Tilde{\algebraicGroup{T}}} q^{-nm/2}\sum_{t \in \tilde{\algebraicGroup{T}}^{\Frobenius}} \tilde{\theta}^{-1}(t)\chi^{-1}({\det}_{\finiteFieldExtension{m}}t) \psi(\trace_{\finiteFieldExtension{m}}t).
\]
Here $\chi \colon \multiplicativegroup{\finiteFieldExtension{m}} \to \multiplicativegroup{\cComplex}$ and $\fieldCharacter \colon \finiteFieldExtension{m} \to \multiplicativegroup{\cComplex}$ are characters. The following lemma shows that this is well defined.

>>>>

The main result of this paper is 
\begin{theorem}
\label{thm:Phi_stable}
	The assignments $H \to \cComplex$ given by $$h \mapsto \genHermitianJacobiKernel{\hermitianSpace}{\chi}\left(h\right)$$ are stable functions. Equivalently, the assignments $$\pi \mapsto \dblJacobiSumScalar{\pi}{\chi}$$ are constant on geometric Lusztig series.
\end{theorem}
This result will be proved in \Cref{subsec:stability}. Our next goal is to prove the following intermediate result.
\begin{lemma}
\label{lem:indep_geo_conj}
    If $(\algebraicGroup{T}_H,\theta_H)$ and $(\algebraicGroup{T}'_H, \theta_H)$ are geometrically conjugate, then $$\frac{\dblVirtualJacobiSumScalar{\RTTheta{T_H}{\theta_H}}{\chi}}{\RTTheta{T_H}{\theta}(1)} = \frac{\dblVirtualJacobiSumScalar{\RTTheta{T'_H}{\theta'}}{\chi}}{\RTTheta{T'_H}{\theta'_H}\left(1\right)}.$$
\end{lemma}
Although will we only use restriction of scalars from $\quadraticExtension$,
we recall in general that $$
	\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N} = \prod_{a \in \mathbb{Z}/m\mathbb{Z}} \left(\algebraicGroup{\GL}_N\right)_a,
$$
with Frobenius root action $$\Frobenius(g_0,g_1, \hdots, g_{m-1}) = (\Frobenius(g_{m-1}), \Frobenius(g_0), \Frobenius(g_1), \hdots, \Frobenius(g_{m-2})).$$ We will freely identify $\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N}(\finiteField) = \GL_N(\finiteFieldExtension{m})$ via the map $\left(g_0,\dots,g_{m-1}\right) \mapsto g_0$.

@@

%Given a rational maximal torus $\algebraicGroup{T} \subset \restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n}$, a character of $\algebraicGroup{T}(\mathbb{F})$, and a non-trivial additive character $\psi$ of $\finiteFieldExtension{m}$, we define a Guass sum
%\[
%	\tau_{\algebraicGroup{T}}(\theta,\psi) := \varepsilon_{\algebraicGroup{
%    \algebraicGroup{T}}}\sum_{t\in \tilde{\algebraicGroup{T}}^{\Frobenius}} \theta^{-1}(t)\psi(\trace(t)).
%\]
%\calvin{maybe should do inverses in the definition. Not sure if $\tau$ or $g$ make more sense.}
Given a torus character pair $(\algebraicGroup{T},\theta)$ with associated $\Frobenius$-twisted conjugacy class $(w) \in W(\algebraicGroup{T}_0)$ along with an embedding $\rho:\algebraicGroup{G}^* \to \restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N}$ and a Frobenius stable maximal torus $\algebraicGroup{T}_{0,\rho} \subset \restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N}$ such that $\rho(\algebraicGroup{T}_0^*) \subset \algebraicGroup{T}_{0,\rho}^*$, let $\tilde{\algebraicGroup{T}}$ be the torus of $\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_N}$ corresponding to the $\Frobenius$-twisted conjugacy class of $\rho(w) \in W(\algebraicGroup{T}_{0,\rho})$ and let $\tilde{\theta} \colon \tilde{\algebraicGroup{T}}^{\Frobenius} \to \multiplicativegroup{\cComplex}$ be a character such that $(\rho(s_{\algebraicGroup{T},\theta})) = (s_{\tilde{\algebraicGroup{T}},\tilde{\theta}})$. We define
\[
    g_{\algebraicGroup{T},\rho}(\chi,\theta,\psi) := \tau_{\Tilde{\algebraicGroup{T}}}(\tilde{\theta} \times \chi,\psi) = \epsilon_{\Tilde{\algebraicGroup{T}}} q^{-nm/2}\sum_{t \in \tilde{\algebraicGroup{T}}^{\Frobenius}} \tilde{\theta}^{-1}(t)\chi^{-1}({\det}_{\finiteFieldExtension{m}}t) \psi(\trace_{\finiteFieldExtension{m}}t).
\]
Here $\chi \colon \multiplicativegroup{\finiteFieldExtension{m}} \to \multiplicativegroup{\cComplex}$ and $\fieldCharacter \colon \finiteFieldExtension{m} \to \multiplicativegroup{\cComplex}$ are characters. The following lemma shows that this is well defined.
